Vagueness, Noise, and Communication

### 2014-10-21 paper/01_intro.tex
conclusion section
Some small changes.

A number of authors have recently tried to explain why vagueness
evolved, based on considerations why a vague language might be useful
\citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,Jaegherde-JaegherRooijvan-Rooij2010:Strategic-Vague,BlumeBoard2013:Intentional-Vag}. In
contrast to these, others have argued that vagueness is a natural
byproduct of limitations in information processing
\citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o,QingFranke2014:Gradable-Adject}. This
paper makes a contribution to the latter line of thought.  Concretely,
we introduce the replicator diffusion dynamics ---a novel variant of
the replicator mutator dynamic--- that integrates stochastic noise on
the differential confusability of similar stimuli. Our main
contribution, therefore, is a technical answer to Lipman's problem in
the sense introduced above. The dynamic proposed here generalizes and
complements previous like-minded accounts. We show that stochastic
noise can not only lead to vague, yet communicative efficient signal
use, but can also unify evolutionary outcomes and help avoid
suboptimal categorization.

@@

The view that vagueness is a natural concomitant of cognitive
limitations of language users has been formalized in a number of ways,
using evolutionary game theory and certain generalizations of
signaling games, so called \emph{similarity-maximizing games}, or
\emph{sim-max games}, for short
\citep{FrankeJager2010:Vagueness-Signa,OConnor2013:Evolving-Percep,OConnor2013:The-Evolution-o}. Our
contribution is best seen in relation to these accounts, as it also
relies on sim-max games. Let's introduce these first, and then zoom in
on the problem of vagueness.

@@

Similarity-maximizing games are generalizations of Lewis' games in
that they allow different states to be more or less similar to one
another, and, roughly put, make success of communication a function of
that similarity. Intuitively speaking, while Lewis' games treated
communicative success as a matter of black and white, sim-max games
allow for a gradient notion of communicative success: the more similar
the receiver's interpretation is to the actual state, the more
successful the play is taken to be. Signaling games with
utility-relevant similarities in the state space are fairly standard
in economics
\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor},
but have received particular attention in a more philosophical and
linguistic context for reasons that will become clear presently
\citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag}.

A sim-max game, in the relevant sense here, consists of a set of
states $\States$, a set of messages $\Messgs$ with much fever messages
than states, a prior probability distribution $\Pr \in
\Delta(\States)$ that gives the occurrence probabilities of states, a
similarity metric on states $\similarity \mycolon \States \times
\States \rightarrow \mathds{R}$, and a utility function $\Utils
\mycolon \States \times \States \rightarrow \mathds{R}$. We identify
the receiver's acts with the states of the world, so that the game is
one of guessing the actual state, so to speak. We also assume that
sender's and receiver's interests are alike, so we only have one
utility function. We do not consider message costs, so utilities only
depend on the actual state, and the receiver's response. The
similarity function should satisfy the usual requirements for a
metric:
\begin{align*}
  & \similarity(\mystate{1}, \mystate{2}) \ge 0 &&
  \similarity(\mystate{1}, \mystate{2}) = 0 \Leftrightarrow
  \mystate{1} = \mystate{2} \\
  & \similarity(\mystate{1}, \mystate{2}) = \similarity(\mystate{2},
  \mystate{1}) && \similarity(\mystate{1}, \mystate{2}) \le
  \similarity(\mystate{1}, \mystate{3}) + \similarity(\mystate{3}, \mystate{2})\,.
\end{align*}
The utility function should be a monotonically decreasing function of
similarity:
\begin{align*}
  \similarity(\mystate{1},\mystate{2}) \ge
  \similarity(\mystate{1},\mystate{3}) \ \ \Rightarrow \ \ 
  \Utils(\mystate{1},\mystate{2}) \ge
  \Utils(\mystate{1},\mystate{3})\,.
\end{align*}
To keep matters simple, we here only look at cases where $\States$
contains finitely many points from the unit interval, all of which
have equal probability, and with similarity given by the Euclidean
distance in this one-dimensional space.

\citet{JagerMetzger2011:Voronoi-Languag} showed that the
evolutionarily stable states of sim-max games are remarkably
systematic. Their results were obtained for games with infinitely many
states in $n$-dimensional Euclidean space $\States \subseteq
\mathds{R}^n$ and a quadratic loss function for utilities
$\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} -
\mystate{2})^2$. For these games, the evolutionarily stable states are
demonstrably so-called Voronoi-languages. Intuitively speaking, a
Voronoi language is a pair of sender and receiver strategies, such
that the sender strategy (quasi-)partitions the state space into
convex categories, while the receiver's interpretations are the
central spots in each category. Figure~\ref{fig:VoronoiLanguage} gives
an example of such a Voronoi-language for a sim-max game with $\States
= [0,1]$, two messages, and a flat prior distribution over states. The
sender uses one message exclusively for all points in the lower half
of the interval and another for all points in the upper half. The
receiver's interpretations of messages are the central points in the
respective intervals.

@@

        \caption{Example of a Voronoi-language on $\States =
          [0,1]$. The (pure) sender strategy $\Spure \in
          \Messgs^\States$ uses one signal for the lower half, and
          another for the upper half of the unit interval. The (pure)
          receiver strategy $\Rpure \in \States^\Messgs$ selects the
          central elements in the respective intervals.}
  \label{fig:VoronoiLanguage}
\end{figure}

This result is interesting, because it demonstrates that signaling can
impose orderly categories on a metric space, without that being the
ulterior purpose of it all. Sender and receiver can be distinct
entities, whose purpose is to communicate effectively about the actual
state. In that case, evolving Voronoi-languages would explain why
linguistic categories are well-behaved and orderly in the way they
appear to be. More abstractly, sender and receiver can also be thought
of as distinct modules in a single system, where the first module must
discretize the information by selecting a small sample of,
suggestively, category labels. These are passed to a second module
that tries to decode the original information. In this case, evolving
Voronoi-languages would explain why conceptual categories are
well-behaved and orderly in the way that they appear to be
\citep[e.g.][for more on this latter
interpretation]{OConnor2013:Evolving-Percep}.

@@

As for micro-dynamics, \citet{FrankeJager2010:Vagueness-Signa} show
how limited memory of past interactions can lead to vague signal use,
when averaging over a single agent's behavior over time or over the
momentary behavior of a population of several language
users. \citet{OConnor2013:The-Evolution-o} introduces a variant of
reinforcement learning that entails a low-level form of stimulus
generalization. Agents update their behavior after each round of play
in such a way that states similar to the one that actually occurred
are subject to behavioral adjustment as well. (A more detailed
discussion of alternative approaches is deferred until
Section~\ref{sec:discussion}, where we can compare them better with the
system introduced in this paper.)

\citet{FrankeJager2010:Vagueness-Signa} also consider a macro-level
approach, using the notion of a \emph{quantal response}. A quantal
response function is a probabilistic choice rule that formalizes the
idea that agents make small mistakes when calculating the expected
utility of choice options. In aggregation, these probabilistic
mistakes lead to systematic ``trembles'' that produce vague signal
use. The approach we take here is superficially similar, but there are
crucial differences. For one, we adopt a dynamic perspective by
looking at limited-time outcomes of a dynamic process. For another, we
demonstrate in Section~\ref{sec:discussion} that quantal responses can
give rise to counterintuitive predictions. These counterintuitive
examples suggest that vagueness in sim-max games is not convincingly
explained by appeal to mistakes in calculating expected utility, but
rather, as we assume here, as the result of confusing similar
states.

>>>>

A number of authors have recently tried to explain why vagueness
evolved, based on considerations why a vague language might be useful
\citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,Jaegherde-JaegherRooijvan-Rooij2010:Strategic-Vague,BlumeBoard2013:Intentional-Vag}. In
contrast to these, others have argued that vagueness is a natural
byproduct of limitations in information processing
\citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o}. This
paper makes a contribution to the latter line of thought.  Concretely,
we introduce the replicator diffusion dynamics ---a novel variant of
the replicator mutator dynamic--- that integrates stochastic noise on
the differential confusability of similar stimuli. Our main
contribution, therefore, is a technical answer to Lipman's problem in
the sense introduced above. The dynamic proposed here generalizes and
complements previous like-minded accounts. We show that stochastic
noise can not only lead to vague, yet communicative efficient signal
use, but can also unify evolutionary outcomes and help avoid
suboptimal categorization.

@@

The view that vagueness is a natural concomitant of cognitive
limitations of language users has been formalized in a number of ways,
using evolutionary game theory and certain generalizations of
signaling games, so called \emph{similarity-maximizing games}, or
\emph{sim-max games}, for short
\citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct}.
%\citep{FrankeJager2010:Vagueness-Signa,OConnor2013:Evolving-Percep,OConnor2013:The-Evolution-o}.
Our contribution is best seen in relation to these accounts, as it also
relies on sim-max games. Let's introduce these first, and then zoom in
on the problem of vagueness.

@@

Similarity-maximizing games are generalizations of Lewis' games where
different states are allowed to be more or less similar to one
another, and, roughly put, success of communication is a function of
that similarity. Intuitively speaking, while Lewis' games treated
communicative success as a matter of black and white, sim-max games
allow for a gradient notion of communicative success: the more similar
the receiver's interpretation is to the actual state, the more
successful the play is taken to be. Signaling games with
utility-relevant similarities in the state space are fairly standard
in economics
\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor},
but have received particular attention in a more philosophical and
linguistic context for reasons that will become clear presently
\citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag}.

A sim-max game, in the relevant sense here, consists of a set of
states $\States$, a set of messages $\Messgs$ with much fever messages
than states, a prior probability distribution $\Pr \in
\Delta(\States)$ that gives the occurrence probabilities of states, a
similarity metric on states $\similarity \mycolon \States \times
\States \rightarrow \mathds{R}$, and a utility function $\Utils
\mycolon \States \times \States \rightarrow \mathds{R}$. We identify
the receiver's acts with the states of the world, so that the game is
one of guessing the actual state, so to speak. We also assume that
sender's and receiver's interests are alike, so we only have one
utility function. We do not consider message costs, so utilities only
depend on the actual state, and the receiver's response. The
similarity function should satisfy the usual requirements for a
metric:
\begin{align*}
  & \similarity(\mystate{1}, \mystate{2}) \ge 0 &&
  \similarity(\mystate{1}, \mystate{2}) = 0 \Leftrightarrow
  \mystate{1} = \mystate{2} \\
  & \similarity(\mystate{1}, \mystate{2}) = \similarity(\mystate{2},
  \mystate{1}) && \similarity(\mystate{1}, \mystate{2}) \le
  \similarity(\mystate{1}, \mystate{3}) + \similarity(\mystate{3}, \mystate{2})\,.
\end{align*}
The utility function should be a monotonically decreasing function of
similarity:
\begin{align*}
  \similarity(\mystate{1},\mystate{2}) \ge
  \similarity(\mystate{1},\mystate{3}) \ \ \Rightarrow \ \ 
  \Utils(\mystate{1},\mystate{2}) \ge
  \Utils(\mystate{1},\mystate{3})\,.
\end{align*}
To keep matters simple, in this paper we mostly focus on cases where $\States$
contains finitely many points from the unit interval, all of which
have equal probability, and with similarity given by the Euclidean
distance in this one-dimensional space.

\citet{JagerMetzger2011:Voronoi-Languag} showed that the
evolutionarily stable states of sim-max games are remarkably
systematic. Their results were obtained for games with infinitely many
states in $n$-dimensional Euclidean space $\States \subseteq
\mathds{R}^n$ and a quadratic loss function for utilities
$\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} -
\mystate{2})^2$. For these games, the evolutionarily stable states are
demonstrably so-called Voronoi languages. Intuitively speaking, a
Voronoi language is a pair of sender and receiver strategies, such
that the sender strategy (quasi-)partitions the state space into
convex categories, while the receiver's interpretations are the
central spots in each category. Figure~\ref{fig:VoronoiLanguage} gives
an example of such a Voronoi language for a sim-max game with $\States
= [0,1]$, two messages, and a flat prior distribution over states. The
sender uses one message exclusively for all points in the lower half
of the interval and another for all points in the upper half. The
receiver's interpretations of messages are the central points in the
respective intervals.

@@

        \caption{Example of a Voronoi language on $\States =
          [0,1]$. The (pure) sender strategy $\Spure \in
          \Messgs^\States$ uses one signal for the lower half, and
          another for the upper half of the unit interval. The (pure)
          receiver strategy $\Rpure \in \States^\Messgs$ selects the
          central elements in the respective intervals.}
  \label{fig:VoronoiLanguage}
\end{figure}

This result is interesting, because it demonstrates that signaling can
impose orderly categories on a metric space, without that being the
ulterior purpose of it all. Sender and receiver can be distinct
entities, whose purpose is to communicate effectively about the actual
state. In that case, evolving Voronoi languages would explain why
linguistic categories are well-behaved and orderly in the way they
appear to be. More abstractly, sender and receiver can also be thought
of as distinct modules in a single system, where the first module must
discretize the information by selecting a small sample of,
suggestively, category labels. These are passed to a second module
that tries to decode the original information. In this case, evolving
Voronoi languages would explain why conceptual categories are
well-behaved and orderly in the way that they appear to be
\citep[e.g.][for more on this latter
interpretation]{OConnor2013:Evolving-Percep}.

@@

As for micro-dynamics, \citet{FrankeJager2010:Vagueness-Signa} show
how limited memory of past interactions can lead to vague signal use,
when averaging over a single agent's behavior over time or over the
momentary behavior of a population of several language
users. \citet{OConnor2013:The-Evolution-o} introduces a variant of
reinforcement learning that entails a low-level form of stimulus
generalization. Agents update their behavior after each round of play
in such a way that states similar to the one that actually occurred
are subject to behavioral adjustment as well.

\citet{FrankeJager2010:Vagueness-Signa} also consider a macro-level
approach, using the notion of a \emph{quantal response}. A quantal
response function is a probabilistic choice rule that formalizes the
idea that agents make small mistakes when calculating the expected
utility of choice options. In aggregation, these probabilistic
mistakes lead to systematic ``trembles'' that produce vague signal
use.

The approach we take here is superficially similar, but there are
crucial differences. For one, we adopt a dynamic perspective by
looking at limited-time outcomes of a dynamic process. For another, we
demonstrate in Section~\ref{sec:discussion} that quantal responses can
give rise to counterintuitive predictions. These counterintuitive
examples suggest that vagueness in sim-max games is not convincingly
explained by appeal to mistakes in calculating expected utility, but
rather, as we assume here, as the result of confusing similar
states. A more in-depth discussion of alternative approaches is deferred until
Section~\ref{sec:discussion}.
Let us first look into our own proposal in more detail.

### 2014-10-21 paper/02_noisyRMD.tex
Some small changes.

Against this background, we introduce a new macro-level approach to
the evolution of vague signal use in sim-max games, where the source
of vagueness is the agents' natural inability to sharply distinguish
similar states. This applies to the sender as well as to
the receiver role (albeit in slightly different manner). We call our
dynamic \emph{replicator diffusion dynamic} (\rdd), because it is a
special case of the replicator mutator dynamic (\rmd).

The \rdd was first introduced by \citet{Correia2013:The-Bivalent-Tr}
as a noise-perturbed variant of the replicator dynamic (\rd) in
behavioral strategies. We begin by recapitulating this formulation of
the \rd in Section~\ref{sec:repl-dynam-behav}. Then we introduce the
\rdd in Section~\ref{sec:repl-diff-dynam-1}.
Section~\ref{sec:repl-mutat-dynam} introduces the \rmd, and
Section~\ref{sec:diffusion-as-special} shows in what sense the \rdd is
a special case of the \rdd.

@@

Fix a signaling game with finite states $\States$, messages $\Messgs$
and acts $\Acts$. Let $\Pr(\cdot) \in \Delta{\States}$ be the prior
distribution over states and $\Util_{\sen,\rec} \mycolon \States
\times \Messgs \times \Acts \rightarrow \mathds{R}$ the senders's and
receiver's utility functions. The sender's behavioral strategies are
functions $\Sstrat \in \Delta(\Messgs)^\States$; the receiver's are
functions $\Rstrat \in \Delta(\Acts)^\Messgs$. Define the expected
utility of choices in each choice point as usual:
\begin{align*}
  \EU(\messg \myts ; \myts \state, \Rstrat) & = \sum_{\act \in \Acts}
  \Rstrat(\act \probbar \messg) \cdot U_\sen(\state, \messg, \act) \\
  \EU(\act \myts ; \myts \messg, \Sstrat) & = \sum_{\state \in
    \States} \Pr(\state) \cdot \Sstrat(\messg \probbar \state) \cdot
  U_\rec(\state, \messg, \act) \,.
\end{align*}
The \emph{fitness} of a behavioral strategy at a choice point is the
frequency-weighted average of expected utilities of each choice, given
the opponent's strategy:
\begin{align*}
  \Phi(\state,\Sstrat, \Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
\EU(\messg \myts ; \myts\state,\Rstrat) \\
\Phi(\messg,\Sstrat, \Rstrat) & = \sum_{\act} \Rstrat(\act \probbar \messg)
\cdot \EU(\act \myts ; \myts \messg,\Sstrat)\,.
\end{align*}
The discrete-time replicator dynamic maps current strategies $\Sstrat$
and $\Rstrat$ to future strategies $\RD(\Sstrat)$ and $\RD(\Rstrat)$
in such a way that changes in frequency are proportional to expected
utilities. For behavioral strategies, the changes take place locally
at each of the agents' choice points:
\begin{align*}
  \RD(\Sstrat)(\messg \probbar \state) & = \frac{\Sstrat(\messg \probbar \state) \cdot
    \EU(\messg \myts ; \myts \state,\Rstrat)} {\Phi(\state,\Sstrat, \Rstrat)} \\
    \RD(\Rstrat)(\act \probbar \messg) & = \frac{\Rstrat(\act \probbar \messg) \cdot
    \EU(\act \myts ; \myts \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
\end{align*}

@@

The replicator diffusion dynamic adds probabilistic confusion of
similar states to the replicator dynamic. Fix a sim-max game with
$\States = \Acts$ and a confusion matrix $C : \States \times
\States$. $C$ is a row-stochastic matrix whose elements $C_{ij}$ give
the probability that $\state_i$ is realized as $\state_j$. The
confusability of states affects senders and receivers alike, but in
slightly different ways (see
Figure~\ref{fig:noise-perturbation-of-strategies}). For the sender,
$C_{ij}$ is the probability that the actual state \mystate{i} is
perceived as \mystate{j}. For the receiver, $C_{ij}$ is the
probability that \mystate{j} is the interpretation that is actually
formed when \mystate{i} is the intended interpretation.

@@

      \begin{pgfonlayer}{background}
        \node [draw=black!50, fill=black!20,fit=(actual) (label)
        (output)] {};
      \end{pgfonlayer}
    \end{scope}

@@

      \begin{pgfonlayer}{background}
        \node [draw=black!50, fill=black!20,fit = (message) 
        (realized) (bla)] {};
      \end{pgfonlayer}
    \end{scope}

@@

The discrete-time replicator diffusion dynamic takes the replicator
dynamic as basic, but factors in the confusion of states at each
update step in a sequential update:
\begin{align*}
  \RDD(\Sstrat) & = \Diff_C(\RD(\Sstrat)) &   \RDD(\Rstrat) & = \Diff_C(\RD(\Rstrat)) \,.
\end{align*}
This is equivalent to the following, perhaps more transparent
formulation, given by \citet{Correia2013:The-Bivalent-Tr}:
\begin{align*}
  \RDD(\Sstrat)(\messg \probbar \state_i) & = \sum_{j} C_{ij} \cdot
  \frac{\Sstrat(\messg \probbar \state_j) \cdot
    \EU(\messg \myts ; \myts \state_j,\Rstrat)}
  {\Phi(\state_j,\Sstrat, \Rstrat)} \\
    \RDD(\Rstrat)(\state_i \probbar \messg) & = \sum_{j} C_{ji} \cdot
  \frac{\Rstrat(\state_j \probbar \messg) \cdot
    \EU(\state_j \myts ; \myts \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
\end{align*}

@@

The \rmd has been proposed first in the context of signaling game
models for the evolution of grammar
\citep[e.g.][]{KomarovaNiyogi2001:The-Evolutionar,NowakKomarova2001:Evolution-of-Un,Nowak2006:Evolutionary-Dy}. It
adds to the replicator dynamic stochastic mutation that captures, in
the context of grammar evolution, the differential learning success of
different dispositions in first language acquisition. The relation of
the \rmd to other prominent evolutionary dynamics is well understood
\citep{PageNowak2002:Unifying-Evolut}. The \rmd has seen further
fruitful applications in the context of signaling games
\citep[e.g.][]{HutteggerSkyrms2010:Evolutionary-Dy,Deo2014:The-Semantic-an}. It
is therefore desirable to relate the \rdd that we propose here to the
\rmd.

@@

Every mixed strategy $\Smixed$ converts to a unique behavioral strategy defined by:
\begin{align}
  \label{eq:CorrespondenceBehavioralMixed}
  \Sstrat(\messg \probbar \state) = \sum_{\Spure(\state) = \messg} \Smixed(\Spure)\,.
\end{align} 
Let $G$ be this mapping from mixed to behavioral strategies. This $G$
is \emph{not} an injection, as many mixed strategies map onto the same
behavioral strategy. In this sense, mixed strategies hold more
information than behavioral strategies about the distribution of pure strategies in a population.

@@

The \rmd is usually given only in its continuous-time form. A
two-population (non-payoff adjusted) formulation is then:
\begin{align*}
  \dot{\Smixed_i} & = \sum_{j}  Q_{ji} \cdot \Smixed_j
    \cdot F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed) &
    \dot{\Rmixed_i} & = \sum_{j}  R_{ji} \cdot \Rmixed_j
    \cdot F_j^{\Smixed} - \Rmixed_i \cdot \Phi(\Rmixed,\Smixed) \,.
\end{align*}
A discrete-time, two-population formulation of the \rmd is this 
\citep[c.f.][97]{PageNowak2002:Unifying-Evolut}:

@@

More concretely, since there is a non-injective mapping $G$ from
behavioral to mixed strategies, we would like to show that there is a
systematic translation from a confusion matrix $C$ to a mutation
matrices $Q^C$ such that for each mixed sender strategy $\Smixed$, its
unique corresponding behavioral strategy $\Sstrat = G(\Smixed)$ has a
confusion-perturbation $C(\Sstrat)$ that corresponds, via $G$, to the
mutation-perturbation $Q^C(\Smixed)$ (see
Figure~\ref{fig:correspondence-result}). In other words, we
hypothesize that a confusion matrix $C$ should give rise to a unique
mutation matrix $Q^C$ so that whenever $G(\Smixed) = \Sstrat$ we also
have $G(Q^C(\Smixed)) = C(\Sstrat)$. Similarly, for the receiver. This
is the content of Theorem~\ref{thm:Correspondence} given below, a
proof for which is in Appendix~\ref{sec:proofs}. The remainder of this
section introduces the relevant construction that maps arbitrary $C$
to suitable $Q^C$.

>>>>

Against this background, we introduce a new macro-level approach to
the evolution of vague signal use in sim-max games, where the source
of vagueness is the agents' natural inability to sharply distinguish
similar states. This applies to the sender as well as to
the receiver role (albeit in slightly different manner). We call our
dynamic \emph{replicator diffusion dynamic} (\rdd), because it is an
extension of the replicator dynamic (\rd) that adds diffusion
of behavior. It is also a special case of the replicator mutator dynamic (\rmd).

We begin by recapitulating the formulation of the \rd in Section~\ref{sec:repl-dynam-behav}.
Then we introduce the \rdd in Section~\ref{sec:repl-diff-dynam-1}.
Section~\ref{sec:repl-mutat-dynam} introduces the \rmd, and
Section~\ref{sec:diffusion-as-special} shows in what sense the \rdd is
a special case of the \rmd.

@@

Fix a signaling game with finite states $\States$, messages $\Messgs$
and acts $\Acts$. Let $\Pr(\cdot) \in \Delta{\States}$ be the prior
distribution over states and $\Util_{\sen,\rec} \mycolon \States
\times \Messgs \times \Acts \rightarrow \mathds{R}$ the senders's and
receiver's utility functions. The sender's behavioral strategies are
functions $\Sstrat \in \Delta(\Messgs)^\States$; the receiver's are
functions $\Rstrat \in \Delta(\Acts)^\Messgs$. Define the expected
utility of choices in each choice point as usual:
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\act \in \Acts}
  \Rstrat(\act \probbar \messg) \cdot U_\sen(\state, \messg, \act) \\
  \EU(\act, \messg, \Sstrat) & = \sum_{\state \in
    \States} \Pr(\state) \cdot \Sstrat(\messg \probbar \state) \cdot
  U_\rec(\state, \messg, \act) \,.
\end{align*}
The \emph{fitness} of a behavioral strategy at a choice point is the
frequency-weighted average of expected utilities of each choice, given
the opponent's strategy:
\begin{align*}
  \Phi(\state,\Sstrat, \Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
\EU(\messg, \state,\Rstrat) \\
\Phi(\messg,\Sstrat, \Rstrat) & = \sum_{\act} \Rstrat(\act \probbar \messg)
\cdot \EU(\act, \messg,\Sstrat)\,.
\end{align*}
The discrete-time replicator dynamic maps current strategies $\Sstrat$
and $\Rstrat$ to future strategies $\RD(\Sstrat)$ and $\RD(\Rstrat)$
in such a way that changes in frequency are proportional to expected
utilities. For behavioral strategies, the changes take place locally
at each of the agents' choice points:
\begin{align*}
  \RD(\Sstrat)(\messg \probbar \state) & = \frac{\Sstrat(\messg \probbar \state) \cdot
    \EU(\messg, \state,\Rstrat)} {\Phi(\state,\Sstrat, \Rstrat)} \\
    \RD(\Rstrat)(\act \probbar \messg) & = \frac{\Rstrat(\act \probbar \messg) \cdot
    \EU(\act, \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
\end{align*}

@@

The replicator diffusion dynamic was first introduced by \citet{Correia2013:The-Bivalent-Tr}
as a noise-perturbed variant of the replicator dynamic (\rd) in
behavioral strategies. Fix a sim-max game with
$\States = \Acts$ and a confusion matrix $C : \States \times
\States$. $C$ is a row-stochastic matrix whose elements $C_{ij}$ give
the probability that $\state_i$ is realized as $\state_j$. The
confusability of states affects senders and receivers alike, but in
slightly different ways (see
Figure~\ref{fig:noise-perturbation-of-strategies}). For the sender,
$C_{ij}$ is the probability that the actual state \mystate{i} is
perceived as \mystate{j}. For the receiver, $C_{ij}$ is the
probability that \mystate{j} is the interpretation that is actually
formed when \mystate{i} is the intended interpretation.

@@

%      \begin{pgfonlayer}{background}
%        \node [draw=black!50, fill=black!20,fit=(actual) (label)
%        (output)] {};
%      \end{pgfonlayer}
    \end{scope}

@@

%      \begin{pgfonlayer}{background}
%        \node [draw=black!50, fill=black!20,fit = (message) 
%        (realized) (bla)] {};
%      \end{pgfonlayer}
    \end{scope}

@@

The discrete-time replicator diffusion dynamic takes the replicator
dynamic as basic, but factors in the confusion of states at each
update step in a sequential update:
\begin{align*}
  \RDD(\Sstrat) & = \Diff_C(\RD(\Sstrat)) &   \RDD(\Rstrat) & = \Diff_C(\RD(\Rstrat)) \,.
\end{align*}
This is equivalent to the following, perhaps more transparent
formulation, given by \citet{Correia2013:The-Bivalent-Tr}:
\begin{align*}
  \RDD(\Sstrat)(\messg \probbar \state_i) & = \sum_{j} C_{ij} \cdot
  \frac{\Sstrat(\messg \probbar \state_j) \cdot
    \EU(\messg, \state_j,\Rstrat)}
  {\Phi(\state_j,\Sstrat, \Rstrat)} \\
    \RDD(\Rstrat)(\state_i \probbar \messg) & = \sum_{j} C_{ji} \cdot
  \frac{\Rstrat(\state_j \probbar \messg) \cdot
    \EU(\state_j, \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
\end{align*}

@@

The \rmd has been proposed first in the context of signaling game
models for the evolution of grammar
\citep[e.g.][]{KomarovaNiyogi2001:The-Evolutionar,NowakKomarova2001:Evolution-of-Un,Nowak2006:Evolutionary-Dy}. It
adds stochastic mutation to the replicator dynamic in order to capture, in
the context of grammar evolution, the differential learning success of
different dispositions in first language acquisition. The relation of
the \rmd to other prominent evolutionary dynamics is well understood
\citep{PageNowak2002:Unifying-Evolut}. The \rmd has seen further
fruitful applications in the context of signaling games
\citep[e.g.][]{HutteggerSkyrms2010:Evolutionary-Dy,Deo2014:The-Semantic-an}. It
is therefore desirable to relate the \rdd that we propose here to the
\rmd.

@@

Every mixed strategy $\Smixed$ converts to a unique behavioral strategy defined by:
\begin{align}
  \label{eq:CorrespondenceBehavioralMixed}
  \Sstrat(\messg \probbar \state) = \sum_{\Spure(\state) = \messg} \Smixed(\Spure)\,.
\end{align} 
Let $G$ be this mapping from mixed to behavioral strategies. Note that $G$
is \emph{not} an injection, as many mixed strategies map onto the same
behavioral strategy. In this sense, mixed strategies hold more
information than behavioral strategies about the distribution of pure strategies in a population.

@@

The \rmd is usually given only in its continuous-time form. A
two-population (non-payoff adjusted) formulation is then:
\begin{align*}
  \dot{\Smixed_i} & = \sum_{j}  Q_{ji} \cdot \Smixed_j
    \cdot F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed) &
    \dot{\Rmixed_i} & = \sum_{j}  R_{ji} \cdot \Rmixed_j
    \cdot F_j^{\Smixed} - \Rmixed_i \cdot \Phi(\Rmixed,\Smixed) \,.
\end{align*}
A discrete-time, two-population formulation of the \rmd can be defined as follows 
\citep[c.f.][97]{PageNowak2002:Unifying-Evolut}:

@@

More concretely, since there is a non-injective mapping $G$ from
behavioral to mixed strategies, we would like to show that there is a
systematic translation from a confusion matrix $C$ to a mutation
matrices $Q^C$ such that for each mixed sender strategy $\Smixed$, its
unique corresponding behavioral strategy $\Sstrat = G(\Smixed)$ has a
confusion-perturbation $C(\Sstrat)$ that corresponds, via $G$, to the
mutation-perturbation $Q^C(\Smixed)$ (see
Figure~\ref{fig:correspondence-result}). In other words, we
hypothesize that a confusion matrix $C$ should give rise to a unique
mutation matrix $Q^C$ so that whenever $G(\Smixed) = \Sstrat$ we also
have $G(Q^C(\Smixed)) = C(\Sstrat)$. Similarly, for the receiver.
%This is the content of Theorem~\ref{thm:Correspondence} given below, a
%proof for which is in Appendix~\ref{sec:proofs}. The remainder of this
%section introduces the relevant construction that maps arbitrary $C$
%to suitable $Q^C$.
